---
title: 'Shiny Tutorial: Precipitation Graphs'
author: "Nathan Byers"
date: "Wednesday, August 20, 2014"
output:
  html_document:
    theme: readable
runtime: shiny
---

This tutorial uses data from the `IDEMdata` package (found on <a href="https://github.com/InDEM/IDEMdata"
target="_blank">GitHub</a>) and explains how to make a Shiny app for  
precipitation graphs (see the multi-part R tutorial found <a href="http://rpubs.com/NateByers/precipRidem"
target="_blank">here</a>, <a href="http://rpubs.com/NateByers/precip2Ridem"
target="_blank">here</a>, and <a href="http://rpubs.com/NateByers/precip3Ridem"
target="_blank">here</a>). 

## Shiny

Shiny is an <a href="http://shiny.rstudio.com/" target="_blank">R package by 
RStudio<a/> that enables you to create interactive graphs. RStudio has a <a 
href="http://shiny.rstudio.com/tutorial/" target="_blank">great tutorial</a> for 
building and deploying Shiny apps. This tutorial won't assume that you've already 
familiar with Shiny, but it's a good idea to utilize RStudio's resources on their 
website.

## Basics

To make a Shiny app you need to have two `.R` scripts in the same folder:

* a user-interface script (ui)
* a server script

## ui.R

The `ui.R` script determines the layout and look of your app. Let's look at the 
`ui.R` for the __Hello Shiny__ example from the RStudio tutorial:

```{r, eval=FALSE}
library(shiny)

# Define UI for application that draws a histogram
shinyUI(fluidPage(

  # Application title
  titlePanel("Hello Shiny!"),

  # Sidebar with a slider input for the number of bins
  sidebarLayout(
    sidebarPanel(
      sliderInput("bins",
                  "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)
    ),

    # Show a plot of the generated distribution
    mainPanel(
      plotOutput("distPlot")
    )
  )
))
```

## server.R

The `server.R` script is where you put the code that will build the app.

```{r, eval=FALSE}

library(shiny)

# Define server logic required to draw a histogram
shinyServer(function(input, output) {

  # Expression that generates a histogram. The expression is
  # wrapped in a call to renderPlot to indicate that:
  #
  #  1) It is "reactive" and therefore should be automatically
  #     re-executed when inputs change
  #  2) Its output type is a plot

  output$distPlot <- renderPlot({
    x    <- faithful[, 2]  # Old Faithful Geyser data
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    # draw the histogram with the specified number of bins
    hist(x, breaks = bins, col = 'darkgray', border = 'white')
  })

})
```

## Running the app locally

To run an app locally, use the `runApp()` function. The `ui.R` and `server.R`
scripts for the __Hello Shiny__ app are found in the `shiny` library, so you can
run this app by placing the directory in the first argument:

```{r, eval=FALSE}
folder <- system.file("examples/01_hello", package="shiny")

runApp(appDir = folder)
```

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/01_hello", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

## Precipitation app

For the precipitation graph Shiny app, we need to first install the latest version
of the IDEMdata package.

```{r, eval=FALSE}
library(devtools)
install_github("InDEM/IDEMdata")
```

Now let's create the `ui.R` file.

```{r, eval=FALSE}
library(shiny)
library(IDEMdata)

data(wide_precip)

# get the station names
stations <- unique(wide.precip.df$station[!is.na(wide.precip.df$station)])

# load the substance_names vector
data(substance_names)

# Define UI for application that draws a histogram
shinyUI(fluidPage(

  # Application title
  titlePanel("Precipitation Graphs"),

  # Sidebar with two dropdown lists
  sidebarLayout(
    sidebarPanel(
      selectInput("station",
                  "Station:", 
                  stations),
      
      selectInput("name",
                  "Sample:", 
                  substance_names)
      
    ),

    # Show the precipitation plot
    mainPanel(
      plotOutput("precipPlot")
    )
  )
))
```

And here's the `server.R` file.

```{r, eval=FALSE}
library(shiny)
library(IDEMdata)
library(lattice)
library(latticeExtra)

data(wide_precip)

# get the sample column names
samples <- colnames(wide.precip.df)[8:34]

# load the substance_names vector
data(substance_names)

# Define server logic required to draw a lattice graph
shinyServer(function(input, output) {

  output$precipPlot <- renderPlot({
    sample.column <- samples[substance_names == input$name]
    
    prSubsetPlot(data = wide.precip.df, site = input$station, site.col = "station", 
                 date.col = "date", precip.col = "precipitation",  
                 sample.col = sample.column, sample.name = input$name, 
                 conditioned = "TRUE", save = FALSE)
  })

})
```

To run the app, we place the `ui.R` and `server.R` scripts in one file and use 
the `runApp()` function. For example, if you place the scripts in a folder called
"precipApp" in your working directory, you can use the following code to open
the app:

```{r, eval=FALSE}
runApp(appDir = "precipApp")
```

```{r, echo=FALSE}
shinyAppDir(
  "precipApp",
  options=list(
    width="100%", height=550
  )
)
```

