---
title: 'Shiny Tutorial: Precipitation Graphs'
author: "Nathan Byers"
date: "Wednesday, August 20, 2014"
output:
  html_document:
    theme: readable
runtime: shiny
---

This tutorial uses data from the `IDEMdata` package (found on <a href="https://github.com/InDEM/IDEMdata"
target="_blank">GitHub</a>) and explains how to make a Shiny app for  
precipitation graphs (see the multi-part R tutorial found <a href="http://rpubs.com/NateByers/precipRidem"
target="_blank">here</a>, <a href="http://rpubs.com/NateByers/precip2Ridem"
target="_blank">here</a>, and <a href="http://rpubs.com/NateByers/precip3Ridem"
target="_blank">here</a>). 

## Shiny

Shiny is an <a href="http://shiny.rstudio.com/" target="_blank">R package by 
RStudio<a/> that enables you to create interactive graphs. RStudio has a <a 
href="http://shiny.rstudio.com/tutorial/" target="_blank">great tutorial</a> for 
building and deploying Shiny apps. This tutorial won't assume that you've already 
familiar with Shiny, but it's a good idea to utilize RStudio's resources on their 
website.

## Basics



```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



